---
title: "rjmcmc_normal"
output: html_document
---

## data

```{r}
source("../dir_manager.r")
source(file.path(dir_path,"rjmcmc_normal/rjmcmc.r"))

set.seed(6291117)
n <-  200
# Generate Data
pi.1 <- 2/10
pi.2 <- 3/10
pi.3 <- 5/10

k <- 3
sim_pi <- c(pi.1, pi.2, pi.3)

sim_mu <- c( -5,    0,    5)
sim_tau <- c(2, 5, 1)
sim_sigma <- 1/sqrt(sim_tau)

y <- c()
for(i in c(1:k)){
  y.k <- rnorm(n*sim_pi[i], mean = sim_mu[i], sd = sim_sigma[i])
  cat("simulated mu k= ",i," :", mean(y.k), "\n")
  cat("simulated tau k= ",i," :", 1/sd(y.k)^2, "\n")
  y <- c(y, y.k)
}

X <- sample(y)

# prior
xi <- 0
kappa <- 0.001
a <- 0.1
g <- 0.1
h <- 1
delta <- 1
lambda <- 3
```


### Birth Death Rate

```{r}
source("../dir_manager.r")
source(file.path(dir_path,"rjmcmc_common/birth_death_prob.r"))

max.K <- 10
min.K <- 1
c_val <- 0.5

res <- get_birth_death_pr(max.K, min.K, lambda, c_val)
birth_pr <- res$birth_pr
death_pr <- res$death_pr
```


### Run

```{r}
source(file.path(dir_path,"rjmcmc_normal/rjmcmc.r"))
set.seed(6291117)
iter <- 100000
burn.in <- iter/5

split_accept_count <- 0
combine_accept_count <- 0
birth_accept_count <- 0
death_accept_count <- 0

split_count <- 0
combine_count <- 0
birth_count <- 0
death_count <- 0

# initial value
k <- 5
b <- rgamma(1, g, h)
mu <- sort(rnorm(k, xi, kappa))
tau <- rep(1, k)
pi <- rdirichlet(1, rep(delta, k))
Z <- rep(1:k, length.out=length(X))
tic("Running RJMCMC normal:")
est <- rjmcmc_mixture_normal_random_b(
  iter, X, max.K,# iteration count, data
  xi, kappa, a, g, h, lambda, delta,#  prior parameters
  k, mu, tau, pi, Z, b, # initial value
  birth_pr, death_pr     )
toc()

est1 <- est[[1]]
lst_est1 <- extract_data(est1, burn.in + 1, iter)

all_Z <- est[[2]]
all_Z <- all_Z[(burn.in + 1): iter,]
```

```{r}
source(file.path(dir_path,"rjmcmc_common/utilities.r"))
source(file.path(dir_path,"rjmcmc_common/rjmcmc_plot.r"))


plot_k_trace_and_density(lst_est1)

est_k <- 3

lst_only3 <- remove_all_k_but(lst_est1, est_k)

```

```{r}
source(file.path(dir_path,"rjmcmc_common/rjmcmc_plot.r"))

# plot_rjmcmc_trace_and_density(lst_only3$mu,  "mu", est_k, sim_mu)
# plot_rjmcmc_trace_and_density(lst_only3$tau,  "tau", est_k, sim_tau)
# plot_rjmcmc_trace_and_density(lst_only3$pi,  "pi", est_k, sim_pi)
pp <- plot_rjmcmc_sixpanel(
  mats   = list(mu = lst_only3$mu, tau = lst_only3$tau, pi = lst_only3$pi),
  k      = est_k,
  truths = list(mu = sim_mu, tau = sim_tau, pi = sim_pi),
  trace_alpha = 0.25,
  legend_position = "bottom",
  panel_label_size = 10
)
ggsave(
  file.path(pic_path, "rjmcmc_normal", "rjmcmc", "raw_3", "raw_3.png"),
  pp,
  width = 8,
  height = 10
)

```


```{r}
source(file.path(dir_path,"rjmcmc_common/post_process.r"))

lst_est_after <- filter_by_component_size(all_Z, lst_est1, min_count = 2)

lst_only3_after <- remove_all_k_but(lst_est_after, est_k)

plot_rjmcmc_sixpanel(
  mats   = list(mu = lst_only3$mu, tau = lst_only3$tau, pi = lst_only3$pi),
  k      = est_k,
  truths = list(mu = sim_mu, tau = sim_tau, pi = sim_pi),
  trace_alpha = 0.25,
  legend_position = "bottom",
  panel_label_size = 10
)
```


# ── Load the data ───────────────────────────────────────────────────────────────

# ── Specify prior and RJMCMC tuning ─────────────────────────────────────────────
GalaxyPrior <- list(
  priorK   = "tpoisson",
  Kmax     = 10,
  delta    = delta,
  priormuQ = "independentC",
  xi       = xi,
  D        = 1/(kappa*10000),
  zeta     = 2 * a,
  g        = g,
  h        = h,
  lambda   = lambda
)

parRJMCMC <- list(
  par.u1 = c(2, 2),
  par.u2 = c(2, 2),
  par.u3 = c(1, 1)
)

# ── Run MCMC ────────────────────────────────────────────────────────────────────
GalaxyModel <- NMixMCMC(
  y0     = X,
  prior  = GalaxyPrior,
  RJMCMC = parRJMCMC,
  nMCMC  = c(burn = 50000, keep = 50000, thin = 1, info = 10000),
  scale  = list(shift = 0, scale = 1),
  PED    = FALSE
)

# ── Basic posterior summary & predictive density ────────────────────────────────
print(GalaxyModel)

```


```{r}
K <- GalaxyModel$K           # Markov chain for K
n_change <- sum(diff(K) != 0)
n_up     <- sum(diff(K) == +1)  # roughly counts birth or split moves
n_down   <- sum(diff(K) == -1) 
```







## second data

```{r}
source("../dir_manager.r")
source(file.path(dir_path,"rjmcmc_normal/rjmcmc.r"))

set.seed(6291117)
n <-  100
# Generate Data
set.seed(123)
pi.1 <- 1/2
pi.2 <- 1/6
pi.3 <- 1/3

sim_pi <- c(pi.1, pi.2, pi.3)

sim_mu <- c( 0,    5,    10)
sim_tau <- c(1/0.5^2, 1/0.5^2, 1/2.5^2)

X.1 <- rnorm(pi.1 * n, sim_mu[1], 1/sqrt(sim_tau[1]) )
X.2 <- rnorm(pi.2 * n, sim_mu[2], 1/sqrt(sim_tau[2]) )
X.3 <- rnorm(pi.3 * n, sim_mu[3], 1/sqrt(sim_tau[3]) )

X <- c(X.1,X.2,X.3)
index <- sample(length(X))
X <- X[index]

# prior
xi <- 0
phi <- 5^2
kappa <- 1/phi
a <- 1
b <- 1
delta <- 1
lambda <- 3
```


### Birth Death Rate

```{r}
source("../dir_manager.r")
source(file.path(dir_path,"rjmcmc_common/birth_death_prob.r"))

max.K <- 10
min.K <- 1
c_val <- 0.5

res <- get_birth_death_pr(max.K, min.K, lambda, c_val)
birth_pr <- res$birth_pr
death_pr <- res$death_pr
```


### Run

```{r}
source(file.path(dir_path,"rjmcmc_normal/rjmcmc.r"))
set.seed(6291117)
iter <- 1000
burn.in <- iter/5

split_accept_count <- 0
combine_accept_count <- 0
birth_accept_count <- 0
death_accept_count <- 0

# initial value
k <- 5
mu <- sort(rnorm(k, xi, kappa))
tau <- rep(1, k)
pi <- rdirichlet(1, rep(delta, k))
Z <- rep(1:k, length.out=length(X))

est <- rjmcmc_mixture_normal(
  iter, X, max.K,# iteration count, data
  xi, kappa, a, b, lambda, delta,#  prior parameters
  k, mu, tau, pi, Z, # initial value
  birth_pr, death_pr     )
est1 <- est[[1]]
lst_est1 <- extract_data(est1, burn.in + 1, iter)

all_Z <- est[[2]]
all_Z <- all_Z[(burn.in + 1): iter,]
```

```{r}
source(file.path(dir_path,"rjmcmc_common/utilities.r"))
source(file.path(dir_path,"rjmcmc_common/rjmcmc_plot.r"))


plot_k_trace_and_density(lst_est1)

est_k <- 3

lst_only3 <- remove_all_k_but(lst_est1, est_k)

```

```{r}
source(file.path(dir_path,"rjmcmc_common/rjmcmc_plot.r"))

plot_rjmcmc_trace_and_density(lst_only3$mu,  "mu", est_k, sim_mu)
plot_rjmcmc_trace_and_density(lst_only3$tau,  "tau", est_k, sim_tau)
plot_rjmcmc_trace_and_density(lst_only3$pi,  "pi", est_k, sim_pi)
```


```{r}
source(file.path(dir_path,"rjmcmc_common/post_process.r"))

lst_est_after <- filter_by_component_size(all_Z, lst_est1, min_count = 5)

lst_only3_after <- remove_all_k_but(lst_est_after, est_k)

plot_rjmcmc_trace_and_density(lst_only3_after$mu,  "mu", est_k, sim_mu)
plot_rjmcmc_trace_and_density(lst_only3_after$tau,  "tau", est_k, sim_tau)
plot_rjmcmc_trace_and_density(lst_only3_after$pi,  "pi", est_k, sim_pi)
```


## Galaxy

```{r}
source("../dir_manager.r")
source(file.path(dir_path,"rjmcmc_normal/rjmcmc.r"))


y <- Galaxy

# prior
R=25.11
g = 0.2
h = 0.016

# prior
xi <- 21.73
kappa <- 1/(R/5)^2
a <- 2
b <- 0.4/0.016
delta <- 1
lambda <- 3
```


# ── Load the data ───────────────────────────────────────────────────────────────

# ── Specify prior and RJMCMC tuning ─────────────────────────────────────────────
GalaxyPrior <- list(
  priorK   = "uniform",
  Kmax     = 30,
  delta    = 1,
  priormuQ = "independentC",
  xi       = 21.73,
  D        = 1/kappa,
  zeta     = 2 * a,
  g        = 0.2,
  h        = 0.016
)

parRJMCMC <- list(
  par.u1 = c(2, 2),
  par.u2 = c(2, 2),
  par.u3 = c(1, 1)
)

# ── Run MCMC ────────────────────────────────────────────────────────────────────
GalaxyModel <- NMixMCMC(
  y0     = Galaxy,
  prior  = GalaxyPrior,
  RJMCMC = parRJMCMC,
  nMCMC  = c(burn = 100000, keep = 100000, thin = 10, info = 10000),
  scale  = list(shift = 0, scale = 1),
  PED    = FALSE
)

# ── Basic posterior summary & predictive density ────────────────────────────────
print(GalaxyModel)

```


### AKPaper mixak



# prior
R = 25.11
g = 0.2
h = 0.016

# prior
xi <- 21.73
kappa <- 0.0016
a <- 2
delta <- 1
lambda <- 3
```








###  RJMCMC

#### Birth Death Rate

```{r}
source("../dir_manager.r")
source(file.path(dir_path,"rjmcmc_common/birth_death_prob.r"))

max.K <- 30
min.K <- 1
c_val <- 0.5

res <- get_birth_death_pr(max.K, min.K, lambda, c_val)
birth_pr <- res$birth_pr
death_pr <- res$death_pr
```


#### Run

```{r}
source(file.path(dir_path,"rjmcmc_normal/rjmcmc.r"))
set.seed(6291117)
iter <- 500000
burn.in <- iter/5

split_accept_count <- 0
combine_accept_count <- 0
birth_accept_count <- 0
death_accept_count <- 0

# initial value
k <- 5
mu <- sort(rnorm(k, xi, kappa))
tau <- rep(1, k)
pi <- rdirichlet(1, rep(delta, k))
Z <- rep(1:k, length.out=length(y))

est <- rjmcmc_mixture_normal(
  iter, y, max.K,# iteration count, data
  xi, kappa, a, b, lambda, delta,#  prior parameters
  k, mu, tau, pi, Z, # initial value
  birth_pr, death_pr     )
est1 <- est[[1]]
lst_est1 <- extract_data(est1, burn.in + 1, iter)

all_Z <- est[[2]]
all_Z <- all_Z[(burn.in + 1): iter,]
```

```{r}
source(file.path(dir_path,"rjmcmc_common/utilities.r"))
source(file.path(dir_path,"rjmcmc_common/rjmcmc_plot.r"))


plot_k_trace_and_density(lst_est1)

est_k <- 4

lst_only3 <- remove_all_k_but(lst_est1, est_k)

```

```{r}

sim_pi <- c(pi.1, pi.2, pi.3)

sim_mu <- c( -10,    0,    8)
sim_tau <- c(15, 10, 30)
source(file.path(dir_path,"rjmcmc_common/rjmcmc_plot.r"))

plot_rjmcmc_trace_and_density(lst_only3$mu,  "mu", est_k, sim_mu)
plot_rjmcmc_trace_and_density(lst_only3$tau,  "tau", est_k, sim_tau)
plot_rjmcmc_trace_and_density(lst_only3$pi,  "pi", est_k, sim_pi)
```


```{r}
source(file.path(dir_path,"rjmcmc_common/post_process.r"))

lst_est_after <- filter_by_component_size(all_Z, lst_est1, min_count = 5)

lst_only3_after <- remove_all_k_but(lst_est_after, est_k)

plot_rjmcmc_trace_and_density(lst_only3_after$mu,  "mu", est_k, sim_mu)
plot_rjmcmc_trace_and_density(lst_only3_after$tau,  "tau", est_k, sim_tau)
plot_rjmcmc_trace_and_density(lst_only3_after$pi,  "pi", est_k, sim_pi)
```




### RJMCMC - gh, compared with poisson AK

#### Birth Death Rate


```{r}


y <- Galaxy
# prior
R = 25.11
g = 0.2
h = 0.016

# prior
xi <- 21.73
kappa <- 0.0016
a <- 2
delta <- 1
lambda <- 3
```


```{r}
source("../dir_manager.r")
source(file.path(dir_path,"rjmcmc_common/birth_death_prob.r"))

max.K <- 10
min.K <- 1
c_val <- 0.5

res <- get_birth_death_pr(max.K, min.K, lambda, c_val)
birth_pr <- res$birth_pr
death_pr <- res$death_pr
```


#### Run

```{r}
source(file.path(dir_path,"rjmcmc_normal/rjmcmc.r"))
set.seed(629)
iter <- 50000
burn.in <- iter/5

split_accept_count <- 0
combine_accept_count <- 0
birth_accept_count <- 0
death_accept_count <- 0

# initial value
k <- 5
mu <- sort(rnorm(k, xi, kappa))
tau <- rep(1, k)
pi <- rdirichlet(1, rep(delta, k))
Z <- rep(1:k, length.out=length(y))
b <- rgamma(1, g, h)

est <- rjmcmc_mixture_normal_random_b(
  iter, y, max.K,# iteration count, data
  xi, kappa, a, g, h, lambda, delta,#  prior parameters
  k, mu, tau, pi, Z, b, # initial value
  birth_pr, death_pr     )
est1 <- est[[1]]
lst_est1 <- extract_data(est1, burn.in + 1, iter)

all_Z <- est[[2]]
all_Z <- all_Z[(burn.in + 1): iter,]
```

```{r}
source(file.path(dir_path,"rjmcmc_common/utilities.r"))
source(file.path(dir_path,"rjmcmc_common/rjmcmc_plot.r"))


plot_k_trace_and_density(lst_est1)

est_k <- 4

lst_only3 <- remove_all_k_but(lst_est1, est_k)

```

```{r}

sim_pi <- c(pi.1, pi.2, pi.3)

sim_mu <- c( -10,    0,    8)
sim_tau <- c(15, 10, 30)
source(file.path(dir_path,"rjmcmc_common/rjmcmc_plot.r"))

plot_rjmcmc_trace_and_density(lst_only3$mu,  "mu", est_k, sim_mu)
plot_rjmcmc_trace_and_density(lst_only3$tau,  "tau", est_k, sim_tau)
plot_rjmcmc_trace_and_density(lst_only3$pi,  "pi", est_k, sim_pi)
```


```{r}
source(file.path(dir_path,"rjmcmc_common/post_process.r"))

lst_est_after <- filter_by_component_size(all_Z, lst_est1, min_count = 5)

lst_only3_after <- remove_all_k_but(lst_est_after, est_k)

plot_rjmcmc_trace_and_density(lst_only3_after$mu,  "mu", est_k, sim_mu)
plot_rjmcmc_trace_and_density(lst_only3_after$tau,  "tau", est_k, sim_tau)
plot_rjmcmc_trace_and_density(lst_only3_after$pi,  "pi", est_k, sim_pi)
```


### RG PF

```{r}

y <- Galaxy
# prior
R=25.11
g = 0.2
h = 0.016

# prior
xi <- 21.73
kappa <- 0.0016
a <- 2
delta <- 1
lambda <- 3

b <- 1

phi <- 1/kappa
set.seed(629)
y <- sample(y)
```



```{r}
source(file.path(dir_path,"particle_filter/normal_fearnhead/pf_normal.r"))
set.seed(6291117)

# prior 
alpha = 1
phi = 1/kappa


N <- 3000


tic("Run Fearnhead PF:")
res <- pf_normal(y, N,
                       a, b, xi, phi, alpha,
                       initial_size = 1,
                      verbose = TRUE) 
toc()
particles <- res[[1]]
weights <- res[[2]]


```

```{r}
source(file.path(dir_path, "particle_filter/normal_fearnhead/normal_helper.r"))
mid_list <- merge_same(particles, weights, toPrint=FALSE)
uni_particles <- mid_list$uni_particles
uni_w <- mid_list$uni_w

N_uni_particles <- length(uni_particles)
cat("unique particles: ", N_uni_particles, "\n")

```



```{r}
all_values <- sort(unique(unlist(uni_particles)))

# Initialize a named vector to store weighted counts
weighted_counts <- setNames(numeric(length(all_values)), all_values)

print(weighted_counts)

# Accumulate weighted occurrence counts
for (i in seq_along(uni_particles)) {
  tab <- length(table(uni_particles[[i]]))
  weighted_counts[tab] <- weighted_counts[tab] + uni_w[i]
}

# Print results
print(weighted_counts)

```



### fearnhead PF

```{r}

y <- Galaxy
# prior
R=25.11
g = 0.2
h = 0.016


alpha <- 3
# prior
xi <- 20
kappa <- 1/15^2
a <- 1
b <- 0.3

phi <- 1/kappa
set.seed(629)
y <- sample(y)
```



```{r}
source(file.path(dir_path,"particle_filter/normal_fearnhead/pf_normal.r"))
set.seed(6291117)

# prior 
alpha = 1
phi = 1/kappa


N <- 300


tic("Run Fearnhead PF:")
res <- pf_normal(y, N,
                       a, b, xi, phi, alpha,
                       initial_size = 1,
                      verbose = TRUE) 
toc()
particles <- res[[1]]
weights <- res[[2]]
```


```{r}
source(file.path(dir_path, "particle_filter/normal_fearnhead/normal_helper.r"))
mid_list <- merge_same(particles, weights, toPrint=FALSE)
uni_particles <- mid_list$uni_particles
uni_w <- mid_list$uni_w

N_uni_particles <- length(uni_particles)
cat("unique particles: ", N_uni_particles, "\n")

```
```{r}
all_values <- sort(unique(unlist(uni_particles)))

# Initialize a named vector to store weighted counts
weighted_counts <- setNames(numeric(length(all_values)), all_values)

print(weighted_counts)

# Accumulate weighted occurrence counts
for (i in seq_along(uni_particles)) {
  tab <- length(table(uni_particles[[i]]))
  weighted_counts[tab] <- weighted_counts[tab] + uni_w[i]
}

# Print results
print(weighted_counts)
```



## Mail

Kia ora Elena, Philipp,

This is an update about the topics we discussed last Tuesday.

Regarding the prior distribution of k and weight in DPMM (or MDP) when the number of observations n is given, it turns out they are not Poisson and Dirichlet. Please see the attachment, where I write down the formulations of FMM and DPMM briefly; based on them, derivations for the distribution of k and weight are presented. So in this case, RJMCMC and Particle Filter are used on different models.

Another important thing is that Richardson & Green (1997) discussed two hierarchical formulations. The first one is called "fixed beta", which I was using. Notation: y_i~N(mu_i, tau_i), and tau_i~Gamma(alpha, beta). Another one is "random beta", where they add an additional hierarchical level on beta, that beta~Gamma(g, h). We discussed at the very beginning that we would have a try for "fixed beta" firstly since I didn't understand why they introduce "random beta". Now, I get the point. They demonstrated that using "fixed beta" cannot achieve "real weak beta prior" for number of clusters k, but "random beta" can (see Fig 6 in R&G 1997). So they recommended the "random beta".


R&G shows the posterior of k for Galaxy data (Table 1) using uniform prior for k, and the distribution being (I ignore those being too small):
    4      5      6      7       8
    0.128  0.182  0.199  0.160   0.109

    4        5        6        7        8
    0.114628 0.215102 0.246970 0.185422 0.108132




mixAk:
    4       5       6       7       8        
    0.17777 0.30451 0.26318 0.13620 0.05062

mine:
    4          5          6          7          8
    0.14721250 0.22689375 0.22128750 0.16322500 0.09778750
    
All iterations are 200,000.


And also, I also tried it on my Fearnhead code, but I can't use same prior because I don't consider the additional level beta~Gamma(g,h), so I just set beta = 1. 
Fearnhead:
  3            4            5            6            7
  1.878340e-01 1.283628e-01 1.327368e-01 4.396580e-01 1.077744e-01

I am using particle number N=3000, which is quite large; but still some valuable particles have been discard during resampling with only part of observations. I think it is the reason why we observe bimodal posterior of k.


I'd like to know how these difference look to you from your experiences; but at least they are way more close compared with the one when I am using "fixed beta" version code. Also I am writing the part in thesis for FMM and normal RJMCMC; as planed, I will check the code with the mathematical terms carefully this week.

